### 2017-04-19 Simulator/article.tex
la2pdf commit - 2017:04:19	21:11:29	(on.19.april)

\usepackage{fixltx2e}                           % Fix some twocolumn float problems

@@

% NOTE: PDF hyperlink and bookmark features are not required in IEEE
%       papers and their use requires extra complexity and work.
% \newcommand\MYhyperrefoptions{bookmarks=true,bookmarksnumbered=true,
% pdfpagemode={UseOutlines},plainpages=false,pdfpagelabels=true,
% colorlinks=true,linkcolor={black},citecolor={black},urlcolor={black},
% pdftitle={Optimal Window Design for the Low Complexity Adaptive Beamformer in Active Sonar Imaging},
% pdfsubject={},
% pdfauthor={Jo Inge Buskenes},
% pdfkeywords={adaptive beamforming, beamforming, complexity, sonar, active}}%
\ifCLASSINFOpdf
%    \usepackage[\MYhyperrefoptions,pdftex]{hyperref}
   \usepackage[pdftex,
    final,                %Keep hyperlink stuff even when in draft mode
    breaklinks=true,      %Break links when necessary
    linktocpage=true,     %Enable link to page?
    linkcolor=black,  %Colour of links to labels within document
    citecolor=black,      %Colour of links to the biliography
    filecolor=red,        %Colour of links to local files
    pagecolor=black,        %Colour of links to other pages withing document
    urlcolor=Blue,        %Colour of the links to external URLs
    colorlinks=true,      %??
    plainpages=false,     %Store roman/arabic numbering differently to avoid
    bookmarksnumbered ]{hyperref}

\else
   \usepackage[\MYhyperrefoptions,breaklinks=true,dvips]{hyperref}
   \usepackage{breakurl}                        % Allows 'dvips' driver to break links
\fi

@@

\newcommand*\A{\mat A}
\newcommand*\B{\mat B}
\newcommand*\C{\mat C}
\newcommand*\E{\mat E}
\renewcommand*\P{\mat P}
\newcommand*\eP{\mat{\hat P}}
\newcommand*\R{\mat R}
\newcommand*\Ri{\R^{-1}}
\newcommand*\eR{\mat{\hat R}}
\newcommand*\eRi{\hat{\mat R}\;\!^{-1}}
\newcommand*\Navg{N_\text{avg}}
\newcommand*\W{\mat W}
\newcommand*\X{\mat X}
\newcommand*\Xd{\X_{\!\Delta}}
\newcommand*\Y{\mat Y}

@@

\renewcommand*\L{\mat \Lambda}
\newcommand*\U{\mat U}
% \renewcommand*\t{\mathtiny{^T}}
% \newcommand*\h{\mathtiny{^H}}
\renewcommand*\t{^T}
\newcommand*\h{^H}

@@

\begin{figure*}[t]\centering%
\graphicsAI[width=\linewidth]{gfx/scene_coordinate_system.svg}%
\caption{Simulator principle. A model of the seafloor and object is loaded into OpenGL. All points in the imaged seafloor region that are visible from the sonar are orthogonally projected onto a view plane. }\label{simulator_coordinate_system}%
\end{figure*}

% \begin{figure}[t!]\centering%
% \graphicsAI[drawing,width=\linewidth]{gfx/specs.svg}%
% \caption{Simulator specifications.}\label{specs}%
% \end{figure}

@@

@@

Before we can render anything with OpenGL we first have to set up the scene. This involves loading the models and then placing them where we want them with the proper orientation.

Our models are stored in regular 3D model files. The simulator loads these with the Open Asset Import library (\texttt{assimp}). This is a portable open source library that supports loading a wide range of 3D model formats in a uniform manner. It converts the model into a node tree where each node contains data such as vertices, facets and textures, all formatted in an OpenGL friendly way.

The next step is to scale, orient and position the objects in the scene, and finally project this scene onto an image plane. This is achieved by applying a set of transformations to each vertex in the model. In OpenGL, the set of transformations can be formalized as
\begin{align*}
\bmat{x\\y\\0\\1}_\text{image} = \P \cdot \V \cdot \M \cdot \bmat{x\\y\\z\\1}_\text{model},
\end{align*}
where $\bmat{\cdot}_\text{model}$ is the loaded model, $\bmat{\cdot}_\text{image}$ is the output image, and $\M$, $\V$ and $\P$ are transformation matrices. These will be described next.

The first transform we apply is the model matrix $\M$, which operates in the model's local coordinate system. It is defined as
\begin{align}
\bmat{x_t \\ y_t \\ z_t \\ w}_\text{\parbox{1cm}{\setlength\baselineskip{0.3cm}transformed\\model}} &= \boldsymbol{T} \cdot \R \cdot \boldsymbol{S} \bmat{x \\ y \\ z \\ w}_\text{model},
\end{align}
where $\boldsymbol{S}, \R$ and $\boldsymbol{T}$ are matrices\todo{be explicit on dependencies?} for scaling, rotation and translation, respectively. These are provided for reference in Appendix \ref{transformation_matrices}.

Next we map the model into world coordinates using the view matrix $\V$. This involves aligning and positioning the model into the view of the camera. In our case the camera represents the phase center of the sonar, and it is placed in the world origin. The transformation take the form of
\begin{align}
\V &= \mat{T}(x_\text{c}, y_\text{c}, z_\text{c}) \cdot \R_\text{x}(\phi_x, 0, 0) \nonumber\\
&= 
\bmat{
1  &  0  &  0  &  x_\text{c} \\
0  &  1  &  0  &  y_\text{c} \\
0  &  0  &  1  &  z_\text{c} \\
0  &  0  &  0  &  1 \\
} \cdot
\bmat{
1  &  0           &  0           &  0 \\
0  &  \cos\phi_x  &  -\sin\phi_x &  0 \\
0  &  \sin\phi_x  &  \cos\phi_x  &  0 \\
0  &  0           &  0           &  1 \\
},
\end{align}
where $(x_\text{c}, y_\text{c}, z_\text{c})$ is the image center coordinates for along track, cross track, and depth, respectively, and $\Phi_x$ is the elevation angle\todo{is this correct?} of the sonar. It is given as
\begin{align*}
\phi_x &= \arctan\left(\frac{y_\text{mean}}{-z_\text{mean}}\right).
\end{align*}\todo{use absolute sign?}

Finally the relevant part of the scene is projected onto the image plane with a projection matrix $\P$. There are two main types of projection, perspective and orthographic. Perspective projection scales all $x$ and $y$ coordinates inversely with the distance to the objects to create a sense of perspective in the image. Orthographic projection, on the other hand, maps all the scene's vertices to a squared cube where each dimension has the range [-1,1]. This will render all objects equally large whether they are in the distance or up close. The matrix that gives orthographic projection is defined as
\begin{align}
\P = \left(\begin{matrix}
\frac{2}{r-l}  &  0              &  0              &  -\frac{r+l}{r-l} \\
0              &  \frac{2}{t-b}  &  0              &  -\frac{t+b}{t-b} \\
0              &  0              &  \frac{-2}{f-n} &  -\frac{f+n}{f-n} \\
0              &  0              &  0              &  1
\end{matrix}\right)
\end{align}
where $l,r,b$ and $t$ are the left, right, bottom and top boundaries of the projection plane, respectively, and where $n$ and $f$ is the near and far limits of the cube that encompass the scene. These parameters need to be adapted to provide a 2D image that matches the image size, which we specify in along track and cross track coordinates ($x,y$). This is straightforward for left and right since matches the $x$-axis
% They define the window through which we observe our scene.
\begin{align}
l &= x_\text{min}  &  r &= x_\text{max}
\end{align}
For the near and far limits one simply needs to make sure the scene will fit in it
\begin{align}
n &= 0\;\text{m}   &  f &= 300\;\text{m} 
\end{align}
How to set the bottom and top bounaries can be inferred from \Fig{camera_top_bottom}. First we define an origin vector that points to the center of the image
\begin{align}
\vec o &= \left[\frac{x_\text{min}+x_\text{max}}{2}, \frac{y_\text{min}+y_\text{max}}{2}, \frac{z_\text{min}+z_\text{max}}{2} \right]\T,
\end{align}
and the distance from the image origin to its $y$-boundaries is given as
\begin{align}
\Delta y &= \frac{y_\text{max} - y_\text{min}}{2}.
\end{align}
To find the nearest and farthest point in the image that will be part of the projected scene, $\alpha$ and $\beta$, we must consider the seafloor rotation angle around the $x$-axis, $\Phi_x$ 
\begin{align}
\alpha &= \vec 0 - \left[0, \Delta y, \Delta y \times \arctan\Phi_\text{x} \right]
\end{align}

@@

\begin{itemize}
\item Kernel tuning.
\item Sharing. PBO. Asynchronous.
\end{itemize}

@@

\begin{itemize}
\item Write to file
\end{itemize}

@@

@@

We have tested the template matching techniqiues on experimental data from the HISAS1030 synthetic aperture sonar (SAS) attached to the HUGIN autonomous underwater vehicle (AUV), both developed by Kongsberg Maritime. Three objects from Jesusbukta (Jesus Bay) in Norway will be presented; two roughly 3\,m long cylinders imaged a couple days apart in Janyary 2009, and a roughly 5\m long torpedo imaged in June 2012.

The resulting classification performance of the template match ing techniques

Data from Jesusbukta

Cylinder 1: 2.9m (short side towards sonar)
Cylinder 2: 3.05m (diagonal)
Torpedo: 5.06m (looks good)

\begin{figure*}[!tbp]\centering%
\subfloat[SAS image of a cylinder with 2.6\;m length and 0.53\;m radius.]{%
\graphicsAI[drawing,width=0.5\linewidth]{gfx/data_cylinder_submerged.svg}%
\label{data_cylinder}}%
\subfloat[Cylinder simulation with parameters estimated from SAS image: Length 2.6\;m, burial depth 0.263\;m, aspect angle 105$^\circ$.]{
\graphicsAI[drawing,width=0.5\linewidth]{gfx/sim_cylinder_submerged_adaptive.svg}%
\label{sim_cylinder}}%
\caption{A SAS image of a cylinder and a template simulation adapted to it.}
\end{figure*}

\newcommand\cdesc[2]{{\raggedright\setlength\fboxsep{0pt}
\fbox{\colorbox[HTML]{#1}{\vrule height8.5pt depth3.5pt width0pt\hspace{.5cm}}}\ \ #2\\}}

@@

\begin{figure*}[!tbp]\centering%
\subfloat[Best template out of a predefined set. Inaccurate cylinder length, alignment and immersion depth. The segmented shadow and highlight regions are misaligned.]{
\graphicsAI[drawing,width=.48\linewidth]{gfx/overlay_cylinder_submerged.svg}\label{overlay_cylinder}%
}%
\hfill%
\subfloat[Adapted template. Template simulated with adapted parameters. Good fit in both highlight and shadow segments.]{%
\graphicsAI[width=.48\linewidth]{gfx/overlay_cylinder_submerged_adaptive.svg}%
}\caption[LoF entry]{\raggedright Simulation image overlaid the segmented SAS image. Segmentation overlay color description:

\parbox{\linewidth}{\small\centering\vspace{.5\baselineskip}
% \newline\centering\newline %\\[.5\baselineskip]
\begin{minipage}{.4\linewidth}
\cdesc{800000}{Template highlight and image highlight}
\cdesc{FF1000}{Template highlight only}
\cdesc{FFEB00}{Image highlight only}
\cdesc{83FF7C}{Background pixels}
\end{minipage}%
\begin{minipage}{.4\linewidth}
\cdesc{000083}{Template shadow and image shadow}
\cdesc{0014FF}{Template shadow only}
\cdesc{00EFFF}{Image shadow only}
\cdesc{00A7FF}{Template shadow and image highlight}
\end{minipage}
}}
% }\label{overlay_cylinder_adaptive}%
% % \vspace{\baselineskip}
% \caption{Segmented simulation image overlaid the segmented SAS image.}
\end{figure*}

>>>>

% \usepackage{fixltx2e}                           % Fix some twocolumn float problems

@@

% NOTE: PDF hyperlink and bookmark features are not required in IEEE
%       papers and their use requires extra complexity and work.
% \newcommand\MYhyperrefoptions{bookmarks=true,bookmarksnumbered=true,
% pdfpagemode={UseOutlines},plainpages=false,pdfpagelabels=true,
% colorlinks=true,linkcolor={black},citecolor={black},urlcolor={black},
% pdftitle={Optimal Window Design for the Low Complexity Adaptive Beamformer in Active Sonar Imaging},
% pdfsubject={},
% pdfauthor={Jo Inge Buskenes},
% pdfkeywords={adaptive beamforming, beamforming, complexity, sonar, active}}%
\ifCLASSINFOpdf
%    \usepackage[\MYhyperrefoptions,pdftex]{hyperref}
   \usepackage[pdftex,
    final,                %Keep hyperlink stuff even when in draft mode
    breaklinks=true,      %Break links when necessary
    linktocpage=true,     %Enable link to page?
    linkcolor=black,  %Colour of links to labels within document
    citecolor=black,      %Colour of links to the biliography
    filecolor=red,        %Colour of links to local files
    pagecolor=black,        %Colour of links to other pages withing document
    urlcolor=Blue,        %Colour of the links to external URLs
    colorlinks=true,      %??
    plainpages=false,     %Store roman/arabic numbering differently to avoid
    bookmarksnumbered ]{hyperref}
% 
% \else
%    \usepackage[\MYhyperrefoptions,breaklinks=true,dvips]{hyperref}%
\fi
%    \usepackage{breakurl}                        % Allows 'dvips' driver to break links

@@

\newcommand*\A{\mat A}
\newcommand*\B{\mat B}
% \newcommand*\C{\mat C}
\newcommand*\E{\mat E}
\renewcommand*\P{\mat P}
\newcommand*\eP{\mat{\hat P}}
\newcommand*\R{\mat R}
\newcommand*\Ri{\R^{-1}}
\newcommand*\eR{\mat{\hat R}}
\newcommand*\eRi{\hat{\mat R}\;\!^{-1}}
\newcommand*\Navg{N_\text{avg}}
\newcommand*\W{\mat W}
\newcommand*\X{\mat X}
\newcommand*\Xd{\X_{\!\Delta}}
\newcommand*\Y{\mat Y}

@@

\renewcommand*\L{\mat \Lambda}
% \newcommand*\U{\mat U}
% \renewcommand*\t{\mathtiny{^T}}
% \newcommand*\h{\mathtiny{^H}}
\renewcommand*\t{^T}
\newcommand*\h{^H}

@@

\begin{figure*}[t]\centering%
\graphicsAI[width=\linewidth]{gfx/scene_coordinate_system.svg}%
\caption{Simulator principle. A model of the seafloor and object is first loaded into OpenGL, and an orthogonal camera view set up to match the desired image region on the seafloor. Then the scene is rendered and an intensity image and a depth map is created. The intensity image along with the depth information is finally combined to form a sonar like image. This is our template.}\label{simulator_coordinate_system}%
\end{figure*}

@@

For lack of a better name, the simulator will be referred to as "FFISim" throughout this article.

Show:
\begin{itemize}
\item Adaptive template matching to avoid large static template libraries,
\item Improved classification performance by feeding simulator with parameters derived off SAS image,
\item Further improved classification by performing a fine-search within a narrow search space around the SAS derived parameters. GPU makes this more powerful.
\end{itemize}

@@

Our models are stored in regular 3D model files. The simulator begins by loading these into a tree of nodes, each representing a vertex in the model with corresponding facet and texture information. Then the vertices are placed into a virtual scene in OpenGL, and a camera set up to capture the part of the scene corresponding to the image. When OpenGL renders the image, it first computes which vertex contributes to which pixel in the image, then computes the intensity and color of the pixel based on the node's facet, texture and scattering model. The code that operates on vertices are called a vertex shader, and the code that operates on pixels are called a fragment shader.

The mapping of the vertices from object coordinates to image coordinates happen through a set of transformations:
%
\begin{align*}
\bmat{x\\y\\0\\1}_\text{image} = \P \cdot \V \cdot \M \cdot \bmat{x\\y\\z\\1}_\text{object},
\end{align*}
%
where $\bmat{\cdot}_\text{model}$ is the loaded model, $\bmat{\cdot}_\text{image}$ is the output image, and $\M$, $\V$ and $\P$ are transformation matrices. These will be described next.

The first transform we apply is the model matrix $\M$, which operates in the model's local coordinate system. It is defined as
%
\begin{align}
\M = \bmat{x_t \\ y_t \\ z_t \\ w}_\text{\parbox{1cm}{\setlength\baselineskip{0.3cm}transformed\\model}} &= \boldsymbol{T} \cdot \R \cdot \boldsymbol{S} \bmat{x \\ y \\ z \\ w}_\text{model},
\end{align}
%
where $\boldsymbol{S}, \R$ and $\boldsymbol{T}$ are matrices\todo{be explicit on dependencies?} for scaling, rotation and translation, respectively. These are provided for reference in Appendix \ref{transformation_matrices}.

Next we map the model into world coordinates using the view matrix $\V$. This involves aligning the camera to precisely fit the image region. In our case the camera represents the phase center of the sonar, and it is placed in the world origin. The transformation then take the form of
%
\begin{align}
\V &= \mat{T}(x_\text{c}, y_\text{c}, z_\text{c}) \cdot \R_\text{x}(\phi_x, 0, 0) \nonumber\\
&= 
\bmat{
1  &  0  &  0  &  x_\text{c} \\
0  &  1  &  0  &  y_\text{c} \\
0  &  0  &  1  &  z_\text{c} \\
0  &  0  &  0  &  1 \\
} \cdot
\bmat{
1  &  0           &  0           &  0 \\
0  &  \cos\phi_x  &  -\sin\phi_x &  0 \\
0  &  \sin\phi_x  &  \cos\phi_x  &  0 \\
0  &  0           &  0           &  1 \\
},
\end{align}
%
where $(x_\text{c}, y_\text{c}, z_\text{c})$ is the image center coordinates for along track, cross track, and depth, respectively, and $\phi_x$ is the elevation angle\todo{is this correct?} of the sonar. It is given as
%
\begin{align*}
\phi_x &= \arctan\left(\frac{y_\text{mean}}{-z_\text{mean}}\right).
\end{align*}\todo{use absolute sign?}

Finally the relevant part of the scene is projected onto the image plane with a projection matrix $\P$. There are two main types of projection, perspective and orthographic. Perspective projection scales all $x$ and $y$ coordinates inversely with the distance to the objects to create a sense of perspective in the image. Orthographic projection, on the other hand, maps all the scene's vertices to a squared cube where each dimension has the range [-1,1]. This will render all objects equally large whether they are in the distance or up close. The matrix that gives orthographic projection is defined as
%
\begin{align}
\P = \left(\begin{matrix}
\frac{2}{r-l}  &  0              &  0              &  -\frac{r+l}{r-l} \\
0              &  \frac{2}{t-b}  &  0              &  -\frac{t+b}{t-b} \\
0              &  0              &  \frac{-2}{f-n} &  -\frac{f+n}{f-n} \\
0              &  0              &  0              &  1
\end{matrix}\right)
\end{align}
%
where $l,r,b$ and $t$ are the left, right, bottom and top boundaries of the projection plane, respectively, and where $n$ and $f$ is the near and far limits of the cube that encompass the scene. These parameters need to be adapted to provide a 2D image that matches the image size, which we specify in along track and cross track coordinates ($x,y$). This is straightforward for left and right since matches the $x$-axis
% They define the window through which we observe our scene.
\begin{align}
l &= x_\text{min}  &  r &= x_\text{max}
\end{align}
%
For the near and far limits one simply needs to make sure the scene will fit in it, say
%
\begin{align}
n &= 0\;\text{m}   &  f &= 300\;\text{m}.
\end{align}
%
The bottom and top boundaries can be inferred from \Fig{camera_top_bottom}. First we define an origin vector that points to the center of the image
%
\begin{align}
\vec o &= \left[\frac{x_\text{min}+x_\text{max}}{2}, \frac{y_\text{min}+y_\text{max}}{2}, \frac{z_\text{min}+z_\text{max}}{2} \right]\T,
\end{align}
%
and the distance from the image origin to its $y$-boundaries is given as
%
\begin{align}
\Delta y &= \frac{|y_\text{max} - y_\text{min}|}{2}.
\end{align}
%
To find the nearest and farthest point in the image that will be part of the projected scene, $\alpha$ and $\beta$, we must consider the seafloor rotation angle around the $x$-axis, $\Phi_x$ 
%
\begin{align}
\vec\alpha &= \vec 0 - \left[0, \Delta y, \Delta y \cdot \tan\Phi_\text{x} \right] \nn
\vec\beta &= \vec 0 + \left[0, \Delta y, \Delta y \cdot \tan\Phi_\text{x} \right]
\end{align}

@@

\todo{
- Kernel tuning.
- Sharing. PBO. Asynchronous.
}

\subsection{Template matching}

To narrow down the search space of possible templates we derive some reliable statistics of the seafloor and object from the SAS image. While FFISim accepts an arbitrarily complex seabed, we find that modeling it as a plane wave with the right tilt works well. The object is more difficult. For instance, we estimate its length and height from the shape of its shadow, but this fails if the object is notably immersed into the seabed at one end.

After the a template is created we decompose it into its highlight and shadow component, and compute a correlation score between these and the corresponding areas in the SAS image. Then the two scores are summed and normalized to form the final correlation score of the template.

@@

\todo{
Write to file
}

%    """
%    
%    stm: Standard template matching, made with SIGMAS.
%    atm: Adaptive template matching, made with FFISim.
%    atm_bury: Adaptive template matching where a search of bury depths is used.
%    atm_bury_restr: Restricted?
%    
%    Adaptiveness made by letting the template matcher use 5 different rotation
%    angles in the ground-range axis. 
%    
%    
%    Data from Jesusbukta
%    
%    Figures are from run 2012.06.16
%    det210 - Cylinder 1: est 2.92m (short side towards sonar)
%    det233 - Cylinder 2: est 3.05m (diagonal)
%    det379 - Torpedo: est 5.06m (looks good)
%    
%    ROC curves from 3 different runs, first two just 2 days apart in 2009, last one in 2012.
%    
%    run_20120616 - last one. Much better with adaptive templates.
%    
%    ROC curves made by average correlation score of shadow/highlight from template matching.
%    
%    
%    fig4b(2015.11.16)
%       Bug that skipped masking of port side (negative y) detections past 180m range fixed
%       
%    fig4 (2015.11.16)
%       Bug leading to incorrect estimation of cylinder orientation fixed.
%          This affects run090601_1. Should be better after this.
%          
%    fig3 (2015.11.09)
%       Detections outside 180m range have been masked (but bug for port side fixed in fig4b).
%       Fixed an incorrectly annotated detection.
%       
%    fig2 (2015.11.02)
%       Height of object estimated from shadow length
%       Let adaptive method try 5 different rotations in ground-range (-4, -2, 0, 2, 4)
%          Improved results, but incorrect classification of:
%          5 detections in first run, and
%          3 detections in second run
%          => For FPR=0.2 the results are slightly worse than with normal adaptive (particularly first run).
%          Images are problematic in these cases:
%             2 detections get wrong orientation because a line (anchor) cross the image
%             Other misclassifications in image regions with poor echo or shadow strength
%          => Not a big issue as not enough time to detect more than ~20% of the detections in a run
%             In a MCM operation these detections would never be used.
%             The runs have 3000 detections each, photographing 20% of these is too much.
%          Performance FPR>0.2 not interesting.
%          Performance FPR<0.2 much better with adaptive template matching.
%          
%    fig (2015.10.23)
%       atm - cylinder orientation and length estimated from SAS image and fed into simulator.
%       
%       ROC curves from last run. Computing the mean correlation score of the shadow and highlight/echo.
%       Early results. Can only check vs. cylinder-like objects.
%       Standard template library created based on estimation of physical size of known cylinder mines.
%       Not possible for adaptive method to perform better when hitting such a target.
%       
%  
%    """

@@

\begin{figure*}[!tbp]\centering%
\graphicsAI[width=\linewidth]{gfx/fig_images_sonar_tagged.svg}\label{fig_image_sonar}%
\caption{SAS images of 2 cylinders and a torpedo located in the Jesus Bay, Norway. Most of the 1000+ detected objects this run are similar to these.}
\end{figure*}

\begin{figure*}[!tbp]\centering%
\graphicsAI[width=\linewidth]{gfx/fig_images_simulator.pdf}\label{fig_image_simulation}%
\caption{SAS images of 2 cylinders and a torpedo located in the Jesus Bay, Norway. Most of the 1000+ detected objects this run are similar to these.}
\end{figure*}

\begin{figure*}[!tbp]\centering%
\graphicsAI[width=\linewidth]{gfx/fig_rocs.pdf}\label{fig_roc_curves}%
\caption{Receiver operating characteristics (ROC) curves.}
\end{figure*}

\begin{figure*}[!tbp]\centering%
\graphicsAI[width=\linewidth]{gfx/fig_overlay.pdf}%
% \parbox{\linewidth}{\small\centering\vspace{.5\baselineskip}
% \newline\centering\newline %\\[.5\baselineskip]
\newcommand\cdesc[2]{{\raggedright\setlength\fboxsep{0pt}
\fbox{\colorbox[HTML]{#1}{\vrule height8.5pt depth3.5pt width0pt\hspace{.5cm}}}\ \ #2\\}}
\begin{minipage}{.4\linewidth}
\cdesc{800000}{Template highlight and image highlight}
\cdesc{FF1000}{Template highlight only}
\cdesc{FFEB00}{Image highlight only}
\cdesc{83FF7C}{Background pixels}
\end{minipage}%
\begin{minipage}{.4\linewidth}
\cdesc{000083}{Template shadow and image shadow}
\cdesc{0014FF}{Template shadow only}
\cdesc{00EFFF}{Image shadow only}
\cdesc{00A7FF}{Template shadow and image highlight}
\end{minipage}%}
\caption{Comparison of classification performance for the standard and adaptive template technique. The three objects were more or less arbitrary picked from the scene. For the roughly 3\,m long cylinders both methods perform similarly, the standard method seemingly with a slight edge. This is no surprise given that the 3\,m long cylinder model that make up the static template database fit the actual object well here. However, for the longer torpedo-like object the adaptive technique is clearly better. This is its merit; for objects and geometries that are not included in the static template database we can expect the adaptive techniques to perform better.}\label{fig_image_simulation}%
\end{figure*}

@@

The performance of our simulator and template matching techniques where measured using experimental data from an interferometric HISAS1030 synthetic aperture sonar (SAS) attached to a HUGIN autonomous underwater vehicle (AUV), both developed by Kongsberg Maritime, Norway. The HISAS1030 is a fully digital phased array with 32 hydrophones, 100\,kHz center frequency and 30\,kHz bandwidth. It is 1.2\,m long, has a half-power beamwidth (HPBW) of 23$\circ$, and is capable of generating synthetic images with a theoretical resolution of 3-4\,cm in both dimensions. The version of HUGIN used also carries an camera that is used for optical inspection of interesting objects.

We present data from three different HUGIN runs from the Jesus Bay in Norway. The first two are taken a couple days apart in January 2009, the last in June 2012. There are roughly 3000 detections in each, most representing by cylinder shaped objects such as barrels, pipe mines and torpedoes. Therefore, the standard template library were created with SIGMAS+ using a basic  cylinder model with a set of geometrical properties commonly found in the scene\todo{Should be more precise. How many templates? How finely were the parameters sampled?}. For a fair comparison, FFISim was restricted to the same cylinder model as well.

In \Fig{fig_image_sonar} we display SAS images of three typical objects in the 2012 run. All are cylindric; the length two first being estimated to roughly 3\,m and the last 5\,m, likely representing cylinder mines and a torpedo, respectively.

Rotation around ground range axis: -4,-2,0,2,4 degrees.

Like the majority of objects in the Jesus Bay they are all cylinders with a length of a few meters. The same objects are simulated in \Fig{fig_image_simulation}, one with SIGMAS and the two other with FFISim. Each of these simulations were based on

The classification performance is quantified with receiver operating characteristics (ROC) curves from all runs in \Fig{fig_roc_curves}. As expected, the 2009 runs proves a greater challenge than that from 2012, with 5 and 3 false positive detections at FPR=20\% in the January 6. and 8. run, respectively. This can partly be explained by incorrect estimation of orientation for a few objects in close proximity to fish trawler tracks, and partly by poor quality in highlight, shadow, or both. However, for FPR<20\% the adaptive techniques do much better than the static template library. For e.g. a mine counter measures (MCM) operation, this is the only region of interest because only a few objects out of the 3000 detections will be visually inspected up close.

- Stm stimulated with length matching test cylinder. Perfect score against this.
- Atm - not given to improve score. Shadow and highlight both need to match, and overfitting could lead to reduced performance.

Good statistics drawn from the image -> adaptive better. Poor statistics -> hit and miss.

1. Comparison with standard template library (ROC enough?)
   ROC: Adaptive better at FPR<0.2 in all cases.
2. Comparison with SIGMAS
   Fairly similar images, performance is similar. 
   Used to create the standard template library.
3. Comparison with search space
   ROC: Additional search always better FPR<0.2.
   Explain search criterion.

SAS images - look good. Random pick objects. Lengths derived from image.
Simulated images - look good. Adaptive simulation more accurate. Otherwise similar. Different background level. Adaptive technique excels for less normal objects/geometries.
Overlap - Provide quantification measure.

1. Adaptive better? Compare with standard. Doing this in ROC
2. Improved classification? ROCs
3. Further improvement by doing fine-search? ROC

An issue the adaptive technique must deal with when estimating object length

As the ROC curves indicate, the adaptive template matching techniques both  ROC curves demonstrate the

The majority of the objects in the Jesus Bay are cylinder shaped. A SAS image of three such objects are displayed in \Fig{fig_image_sonar}

% \begin{figure*}[!tbp]\centering%
% \subfloat[SAS images of 2 cylinders and a torpedo.]{%
% \graphicsAI[width=0.5\linewidth]{gfx/fig_images_sonar_simulation.pdf}%
% \label{fig_sonar_simulation}}%
% \subfloat[Cylinder simulation with parameters estimated from SAS image: Length 2.6\;m, burial depth 0.263\;m, aspect angle 105$^\circ$.]{
% \graphicsAI[drawing,width=0.5\linewidth]{gfx/sim_cylinder_submerged_adaptive.svg}%
% \label{sim_cylinder}}%
% \caption{A SAS image of a cylinder and a template simulation adapted to it.}
% \end{figure*}
% 
% 
% \begin{figure*}[!tbp]\centering%
% \subfloat[SAS image of a cylinder with 2.6\;m length and 0.53\;m radius.]{%
% \graphicsAI[drawing,width=0.5\linewidth]{gfx/data_cylinder_submerged.svg}%
% \label{data_cylinder}}%
% \subfloat[Cylinder simulation with parameters estimated from SAS image: Length 2.6\;m, burial depth 0.263\;m, aspect angle 105$^\circ$.]{
% \graphicsAI[drawing,width=0.5\linewidth]{gfx/sim_cylinder_submerged_adaptive.svg}%
% \label{sim_cylinder}}%
% \caption{A SAS image of a cylinder and a template simulation adapted to it.}
% \end{figure*}

@@

### 2017-04-19 ThesisJo/introduction.tex
la2pdf commit - 2017:04:19	21:11:29	(on.19.april)

Preface?

@@

>>>>

big data. needed to evolve ai intelligence. e.g. deep learning.  need processing speed to handle it. especially near sensors

data now enough? we have all of evolution.

processimg loop small. learn to adapt behaviour then and there.

arrays allow focus and width.but requires processing.

human information 2000 titanics of thumbdrives. ai used to map genes 
Preface?

@@

One of mankind's greatest dreams is that of creating intelligent machines. Imagine things you can strike a conversation with, that aides us in our daily lives, that never grow old, impatient or boring. Yesterday science fiction, tomorrow reality. It is not a question whether the intelligent machines are coming, but when and how. Around the corner are self driving cars, virtual reality, ...

To understand how we can evolve the intelligence of machines, perhaps it is best to look at how nature has done it. Our 5 basic senses are sight, smell, hearing, taste and touch. The information here is converted from the respective physical phenomena to electrical impulses through various cells. Then these electrical signals travel to a vast network of neurons that make up our brain. Since each neuron is linked to several others, the data dependency is many to many and very hard to understand.

Intelligence is by its essence a feedback loop. 
This requires them to sense their surroundings and To do so, we need equip it with sensors we grant them sensory input by converting energy in various physical forms into electrical signals, process these using analog and digital electronics, and imbue them with  Most of the components for doing so

Two big topics in recent years are big data and articifical intelligence.
